---
title: "Data Cleanup"
author: "Robert Hensley"
date: "March 6, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(jsonlite)
library(tidyverse)
library(yelpr)

key <- "pxKTGs3uKpDZXQSo1nKmJKSPycfRqixcudXpKm8oAuVRNEjx_MVUhBryPP8Gg5MgPn6Cl28RO6dfRNFrh3jNQVdcDxu-uNI5IUT7-phwJPiOW2L2D2cRrEtIXJyAXHYx"

```

## Yelp Fusion API

Example of 10 chinese restaurants in New York:

```{r}
# search businesses with keyword 'chinese' in 'New York'
business_ny <- business_search(api_key = key,
                location = 'New York',
                term = "chinese",
                limit = 10)
```

## San Luis Obispo Data

```{r}
business_SLO <- as.data.frame(business_search(api_key = key, location = 'San Luis Obispo', limit = 50))
nrow(business_SLO)
```

It looks like the API limit of restaurants to view is 50 per call. RIP

If only there were a way to combine the data...

## Combining/Cleaning the Data

```{r}
businesses <- business_search(api_key = key, location = 'San Luis Obispo', limit = 50)
num <- (as.data.frame(businesses) %>% select(total))[1,1]

num
```

Notice that the total value of the dataframe is 340. That means that there are 340 businesses in San Luis Obispo with a Yelp page. 327 results show up on a Yelp Search of San Luis Obispo, so it is assumed that the remaining 13 establishments are closed. 

Before the data can be appended, it must be cleaned. There are several nested data frames in the columns that need to be expanded. Below are two sample dataframes the will intend to merge and information about that data.

```{r}
businesses_test <- business_search(api_key = key, location = 'San Luis Obispo', limit = 50)
append <- business_search(api_key = key, location = 'San Luis Obispo', limit = 50, offset = 50)

b <- as.data.frame(businesses_test, row.names = c(1:50))
a <- as.data.frame(append, row.names = c(51:100))

glimpse(a)
```

Information to remove (to allow binding):
- businesses.coordinates
- businesses.location

```{r}
region.center <- b[1,]$region.center.latitude

# remove closed businesses
#b <- b %>%
  #filter(businesses.is_closed == FALSE)

b <- as.data.frame(businesses_test, row.names = c(1:50))
a <- as.data.frame(append, row.names = c(51:100))

# spread dataframe columns
b["latitude"]  <- businesses[["businesses"]][["coordinates"]][["latitude"]]
b["longitude"] <- businesses[["businesses"]][["coordinates"]][["longitude"]]
b["address"] <- businesses[["businesses"]][["location"]][["address1"]]
b["zip_code"] <- businesses[["businesses"]][["location"]][["zip_code"]]
b["state"] <- businesses[["businesses"]][["location"]][["state"]]
b["country"] <- businesses[["businesses"]][["location"]][["country"]]

a["latitude"]  <- append[["businesses"]][["coordinates"]][["latitude"]]
a["longitude"] <- append[["businesses"]][["coordinates"]][["longitude"]]
a["address"] <- append[["businesses"]][["location"]][["address1"]]
a["zip_code"] <- append[["businesses"]][["location"]][["zip_code"]]
a["state"] <- append[["businesses"]][["location"]][["state"]]
a["country"] <- append[["businesses"]][["location"]][["country"]]

combo <- rbind(b %>% select(-c(businesses.coordinates, businesses.location)), a %>% select(-c(businesses.coordinates, businesses.location)))

head(combo)

```

This method is kind of slow, but it works!

## A function for Data Collection

```{r}
get_businesses <- function(my_location) {
  
  key <- "pxKTGs3uKpDZXQSo1nKmJKSPycfRqixcudXpKm8oAuVRNEjx_MVUhBryPP8Gg5MgPn6Cl28RO6dfRNFrh3jNQVdcDxu-uNI5IUT7-phwJPiOW2L2D2cRrEtIXJyAXHYx"
  
  businesses <- business_search(api_key = key, location = my_location, limit = 50)
  
  # create the initial dataframe
  b <- as.data.frame(businesses, row.names = c(1:50))
  b["latitude"]  <- businesses[["businesses"]][["coordinates"]][["latitude"]]
  b["longitude"] <- businesses[["businesses"]][["coordinates"]][["longitude"]]
  b["address"] <- businesses[["businesses"]][["location"]][["address1"]]
  b["zip_code"] <- businesses[["businesses"]][["location"]][["zip_code"]]
  b["state"] <- businesses[["businesses"]][["location"]][["state"]]
  b["country"] <- businesses[["businesses"]][["location"]][["country"]]
  b <- b %>% select(-c(businesses.coordinates, businesses.location))
  
  num <- (b %>% select(total))[1,1]
  my_offset <- 50 # max number of businesses per API call
  
  while(my_offset < num) {
    # create data to append
    append <- business_search(api_key = key, location = my_location, limit = 50, offset = my_offset)
    a <- as.data.frame(append)
    a["latitude"]  <- append[["businesses"]][["coordinates"]][["latitude"]]
    a["longitude"] <- append[["businesses"]][["coordinates"]][["longitude"]]
    a["address"] <- append[["businesses"]][["location"]][["address1"]]
    a["zip_code"] <- append[["businesses"]][["location"]][["zip_code"]]
    a["state"] <- append[["businesses"]][["location"]][["state"]]
    a["country"] <- append[["businesses"]][["location"]][["country"]]
    
    b <- rbind(b, a %>% select(-c(businesses.coordinates, businesses.location)))
    my_offset = my_offset + 50
  }
  
  b
}
```

Now let's give it a shot!

```{r}
businesses_SLO <- get_businesses("San Luis Obispo")
```

```{r}
head(businesses_SLO)
nrow(businesses_SLO)

businesses_SLO %>% 
    toJSON() %>%
    write_lines("businesses_SLO.json")
```

```{r}

# testing categories merge

test <- head(businesses_SLO) %>% select(businesses.id, businesses.categories)

test2 <- head(businesses_SLO)

categories <- businesses_SLO %>% 
  select(businesses.id, businesses.categories) %>%
  unnest() 

categories_matrix <- categories %>%
  spread(key = alias, value = title)

categories_matrix

categories_map <- data.frame(alias = unique(categories$alias), title = unique(categories$title))

categories_map

```

Success! A future goal would be to speed up the data cleaning process to use on cities with larger datasets.


